---
title: "Classified Ads for Cars Dataset Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(knitr)
```
```{r message=FALSE, warning=FALSE}
set.seed(42)
library(ggplot2)
library(reshape2)
library(plyr)
library(readr)
library(fpc)
library(data.table)
library(ggplot2)
```

## Connecting to USED_CARS database in HIVE running on Hortonworks Sandbox VM

```
options( java.parameters = "-Xmx8g" )
library(rJava)
library(RJDBC)
 
cp = c("//d:/tools/apache-hive-1.2.2/lib/hive-jdbc-1.2.2-standalone.jar",
       "//d:/tools/hadoop-2.7.7/share/hadoop/common/hadoop-common-2.7.7.jar")
.jinit(classpath=cp) 
 
drv <- JDBC(driverClass = "org.apache.hive.jdbc.HiveDriver",
            classPath = "//d:/tools/apache-hive-1.2.2/lib/hive-jdbc-1.2.2-standalone.jar",
             identifier.quote="`")
 
conn <- dbConnect(drv, "jdbc:hive2://127.0.0.1:10000/used_cars", "maria_dev", "maria_dev")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
options( java.parameters = "-Xmx8g" )
library(rJava)
library(RJDBC)
 
cp = c("//d:/tools/apache-hive-1.2.2/lib/hive-jdbc-1.2.2-standalone.jar",
       "//d:/tools/hadoop-2.7.7/share/hadoop/common/hadoop-common-2.7.7.jar")
.jinit(classpath=cp) 
 
drv <- JDBC(driverClass = "org.apache.hive.jdbc.HiveDriver",
            classPath = "//d:/tools/apache-hive-1.2.2/lib/hive-jdbc-1.2.2-standalone.jar",
             identifier.quote="`")
 
conn <- dbConnect(drv, "jdbc:hive2://127.0.0.1:10000/used_cars", "maria_dev", "maria_dev")
dbSendUpdate(conn, "USE used_cars")

```

## Discover the USED_CARS database

First we are getting the EVENTS table description. Full description of the attributes presented on [Classified Ads for Cars](https://www.kaggle.com/mirosval/personal-cars-classifieds/home) home page.

```{r}
descr <- dbGetQuery(conn, "describe events")
kable(descr)
```

Code below selecting first 10 rows for EVENTS table and saves them to 'cars' dataframe. Note that we rename cars columns for better presentation. Also note that it looks that the dataset has some missing values.

```{r results='asis'}
cars <- dbGetQuery(conn, "select * from events limit 6")
colnames(cars) <- c(
  "Maker", "Model","Mileage","Year", "Disp", "Pwr", "Body", "Color", "Sticker", 
  "Trans", "Doors", "Seats", "Fuel", "Listed", "Removed", "Price")
```

```{r results='asis'}
kable(cars)
```

The EVENTS table has 3,552,912 records which could be confirmed by running the HQL statement below.

```
dbGetQuery(conn, "select count (*) from events")
```
```{r eval=FALSE, include=FALSE}
# dbGetQuery(conn, "select count (*) from events")
```

For the analysis we will extract not more than 30K rows from EVENT table randomly using the HQL statement below. We skip the rows mising attributes except for the ones not importants and rarely filled like body_type, color_slug and stk_year:

```{r}
filter <- paste(
      " maker <> ''",
      " AND model <> ''",
      " AND mileage is not NULL",
      " AND manufacture_year is not NULL",
      " AND engine_displacement is not NULL",
      " AND engine_power is not NULL",
      " AND transmission <> ''",
      " AND door_count is not NULL",
      " AND seat_count is not NULL",
      " AND fuel_type <> ''",
      " AND date_created is not NULL",
      " AND date_last_seen is not NULL"
) 
count <- dbGetQuery(conn, paste("select count(*) from events", " WHERE", filter))
cars.sample.totalFilered <- count$`_c0`
```


```{r}
car.sample.maxSize <- 30000
lim <- car.sample.maxSize/cars.sample.totalFilered

sample_HQL <- paste(
  "select * from events WHERE", filter,
      " AND rand(123) < ", lim,
  " limit ", 1.2 * car.sample.maxSize
) 

cars.sample <- dbGetQuery(conn, sample_HQL)
colnames(cars.sample) <- c(
  "Maker", "Model","Mileage","Year", "Disp", "Pwr", "Body", "Color", "Sticker", 
  "Trans", "Doors", "Seats", "Fuel", "Listed", "Removed", "Price")

nrow(cars.sample)
```

```{r results='asis'}
kable(head(cars.sample))
```

```{r}
# summary(cars.sample)
```


```{r hist_1, fig.pos = 'h', fig.height=8, fig.width=5.5, fig.align="center", fig.cap="Number of Ads by Maker"}
require(forcats)
total <- nrow(cars.sample)
ggplot(cars.sample, aes(fct_rev(fct_infreq(Maker)))) +
       geom_bar() + 
       geom_text(stat = "count", aes(label = ..count.., y = ..count..), nudge_y = 120)+
#       theme(axis.text.x = element_text(angle =90 , hjust = 1, vjust = 0.5)) +
       labs(x="", y="Percent of Ads in the Sample Set") +
#       scale_y_log10(breaks = c(1, 2, 5, 10, 25, 50, 100, 250, 500, 1000)) +
        scale_y_continuous(labels = function(x) sprintf("%.0f%%",x/total*100)) +
       coord_flip()
  
```

## Disconnecting from the HIVE

It is very important to disconnect from the HIVE at the end of the session:

```{r}
dbDisconnect(conn)
```
